---
title: "Q vs Snow Cover"
author: "Bryce Pulver"
date: "September 10, 2019"
output: html_document
---

###

```{r setup, include=FALSE}
install.packages("dataRetrieval")
library(tidyverse) 
library(lubridate)
library(dataRetrieval) #for downloading USGS data
knitr::opts_chunk$set(echo = TRUE)
```

### Data Checking

## Data read in

First we need to get our snow metric (ndsi) data and then also download discharge data the USGS

```{r read-data}
#ndsi 
ndsi <- read_csv('data/hayman_ndsi.csv') %>%
  rename(burned=2,unburned=3) %>%
  filter(!is.na(burned),
         !is.na(unburned)) %>%
  gather(.,key='site',
         value='ndsi',
         -DateTime) # For this analysis we want the data in long format
```


```{r read-USGSdata}
#USGS gauge above cheeseman lake '00060'
q_hayman <- readNWISdata(sites=c('06700000'), #Site code 
                  parameterCd='00060', #discharge code in cfs
                  service='dv', # service = daily values (versus annual)
                  startDate='1984-10-01', #Start date for getting the data
                  endDate = '2019-9-10') %>% # End date (today) 
  rename(q_cfs = X_00060_00003,
         quality_cd = X_00060_00003_cd) %>% #rename long column name
  filter(!is.na(q_cfs)) %>% #Drop NAs which can occur when there is ice or sensor breaks
  as_tibble() #To make it act like a tibble
```

### Data Exploring

## NDSI summary
```{r NDSIsummary}
summary(ndsi)
knitr::opts_chunk$set(echo = F)
```

## Q summary
```{r Qsummary}
summary(q_hayman)
knitr::opts_chunk$set(echo = F)
```
### Combining the data

## Adding a water year column

When analyzing water flux data, we typically break the year up into “water years” which run from October to the end of September. For this exploratory analysis, we want to group the datasets by water year and then join them to each other so we can compare winter average, max, median, etc… of snow cover versus the next water year’s water flux. So we have to add a column called water year

# Q Water Year
```{r Qwateryear}
q_water_year <- q_hayman %>%
  mutate(month=month(dateTime),
         year_offset = ifelse(month > 9,1,0),
         wtr_yr = year(dateTime) + year_offset)

table(q_water_year$wtr_yr)
```

# NDSI water year

```{r NDSIwateryear}
ndsi_water_year <- ndsi %>%
  mutate(month=month(DateTime),
         year_offset = ifelse(month > 9,1,0),
         wtr_yr = year(DateTime) + year_offset)

table(ndsi_water_year$wtr_yr)
```
### Filtering and Summarizing

Now that we have our matched datasets we want to do a couple filtering operations. First, we want to make sure that we are only analyzing complete water years from the Q dataset. Second, we want to make sure we are only summarizing the snow data over months where snow cover is possible, which I would guess is between october and may. Once we have these filtering operations done, we want to summarize each dataset by water year so we can eventually join them and see if snow cover predicts Q.

# Snow water year summary statistics

```{r NDSIsummarystats,echo=F}
snow_means <- ndsi_water_year %>%
  filter(month %in% c (10:12,1:5)) %>%
  group_by(wtr_yr) %>%
  summarize(mean_ndsi=mean(ndsi),
            max_ndsi=max(ndsi),
            meadian_ndsi=median(ndsi))

snow_means

```

# Whole Q Water Year Summary 
```{r}
q_means <- q_water_year %>%
  group_by(wtr_yr) %>%
  mutate(count = n()) %>%
  ungroup() %>%
  filter(count == max(count)) %>%
  group_by(wtr_yr) %>%
  summarize(mean_q=mean(q_cfs),
            max_q=max(q_cfs),
            median_q=median(q_cfs))
  
q_means
  

```


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
